SU($n$) spin models near the Heisenberg critical point

### 2020-03-25 collective_sun/sun_perturbation.tex
finalized conventions with symmetry factors, etc. in the definition and product of multi-body operators

\newcommand{\A}{\mathcal{A}}
\newcommand{\B}{\mathcal{B}}
\newcommand{\C}{\mathcal{C}}
\newcommand{\D}{\mathcal{D}}
\newcommand{\E}{\mathcal{E}}
\newcommand{\G}{\mathcal{G}}
\newcommand{\I}{\mathcal{I}}
\renewcommand{\L}{\mathcal{L}}
\newcommand{\M}{\mathcal{M}}
\newcommand{\N}{\mathcal{N}}
\renewcommand{\O}{\mathcal{O}}
\renewcommand{\P}{\mathcal{P}}
\renewcommand{\S}{\mathcal{S}}
\newcommand{\T}{\mathcal{T}}
\newcommand{\Z}{\mathcal{Z}}

@@

We consider an array of $N$ multilevel spins with SU($n$)-symmetric
interactions that can be written in the form
\begin{align}
  H_0 = \sum_{\p{p,q}\in\C_N\p{2}} h_{pq} \Pi_{pq},
  &&
  \Pi_{pq} \equiv \sum_{\mu,\nu\in\ZZ_n} s_{\mu\nu}^{(p)} s_{\nu\mu}^{(q)},
  \label{eq:H_0}
\end{align}
where $p,q$ index individual spins; $\C_N\p{M}$ is the set of all
choices of $M$ spins out of $N$; $h_{pq}$ are scalar coupling
constants; $\mu,\nu$ index states in an orthonormal basis for the
$n$-dimensional Hilbert space of a single spin; $\ZZ_N$ is the set of
integers modulo $N$; the operator
$s_{\mu\nu}^{(p)}\equiv\op{\mu}{\nu}_p$ flips the state of spin $p$ to
$\ket\mu$ from $\ket\nu$; and the operator $\Pi_{pq}$ permutes the
states of spins $p$ and $q$.  The permutation operator $\Pi_{pq}$ is
essentially a multilevel generalization of the SU(2)-symmetric spin
interaction $\v s_p\c\v s_q$, with $\v s\equiv\p{\v s_\x,s_\y,s_\z}$,
$s_\alpha\equiv\sigma_\alpha/2$, and $\sigma_\alpha$ a single-spin
Pauli operator.

If the coefficients $h_{pq}$ of the interaction Hamiltonian $H_0$ are
all negative, then the ground-state manifold $\M_0$ of the interaction
Hamiltonian $H_0$ consists of permutationally symmetric (PS) states
that are simultaneous $+1$ eigenstates of all permutation operators
$\Pi_{pq}$.  We will assume that all $h_{pq}<0$ throughout these
notes, but comment that this restriction can be relaxed to the
assumption that the initial state of all spins is permutationally
symmetric (e.g.~a spin-polarized state).  We will also assume that the
PS manifold $\M_0$ is gapped away from all orthogonal states by an
interaction energy $\Delta_{\t{gap}}>0$.  Power-law couplings of the
form $h_{pq}\sim-1/\abs{p-q}^\alpha$ on a $D$-dimensional lattice, for
example, always yield a non-vanishing spectral gap $\Delta_{\t{gap}}$
for long-range interactions with $\alpha\le D$ (see Appendix
\ref{sec:spectral_gap}).

The PS manifold $\M_0$ is spanned by a basis of states
$\ket{m}=\ket{m_1,m_2,\cdots,m_n}$ with a definite occupation number
$m_\mu$ of the single-spin state $\mu$, and $\sum_\mu m_\mu=N$.  In
the case of $n=2$, for example, the PS manifold $\M_0$ is precisely
the Dicke manifold\cite{dicke1954coherence} of $N$ qubits, spanned by
Dicke states $\set{\ket{m_0,m_1}}$ for which $m_0+m_1=N$ and the
integer $m_0$ ($m_1$) indicates the total occupation number of qubit
state $\ket{0}$ ($\ket{1}$).  The dimension of the PS manifold $\M_0$
is determined by the number of ways to assign $N$ spins to $n$ states,
which is ${N+n-1 \choose n-1}\sim N^{n-1}$.  The energy of any PS
state $\ket\psi\in\M_0$ with respect to the interaction Hamiltonian
$H_0$ in \eqref{eq:H_0} is
\begin{align}
  E_0 = \sum_{\p{p,q}\in\C_N\p{2}} h_{pq}.
\end{align}

@@

In order to induce non-trivial dynamics for initially PS states
$\ket\psi\in\M_0$, which are otherwise stationary eigenstates of the
interaction Hamiltonian $H_0$ in \eqref{eq:H_0}, we will add to the
Hamiltonian an operator $\hat\O$ of the general form
\begin{align}
  \hat\O \equiv \sum_{\p{X,\m w}\in\O} X_{\m w},
  &&
  X_{\m w} \equiv \f1{M!} \sum_{k\in\D_N\p{M}} w_k X_k,
  \label{eq:pert}
\end{align}
where $\hat\O$ is constructed from the set $\O\equiv\set{\p{X,\m w}}$
of $M$-spin operators $X$ together with dimension-$M$ (i.e.~$M$-index)
tensors $\m w$; $k=\p{k_1,k_2,\cdots,k_M}$ is a list of of $M$ spins
that $X_k$ acts on; $w_k$ is a scalar entry of $\m w$; and
\begin{align}
  \D_N\p{M} \equiv
  \set{ k \in \ZZ_N^M:~\t{all elements of}~k~\t{are distinct} },
  \label{eq:off_diags}
\end{align}
is the set of all ordered choices of $M$ spins out of $N$.  The
normalization factor of $1/M!$ in \eqref{eq:pert} is included to
account for the fact that there are $M!$ terms addressing any given
choice of $M$ of spins.  Due to the permutational symmetry of the
ground-state manifold $\M_0$ of the interaction Hamiltonian $H_0$, we
will find it convenient to decompose
$\D_N\p{M}=\pi\p{M}\times\C_N\p{M}$, where $\pi\p{M}$ is the
permutation group of $\ZZ_M$, and $\C_N\p{M}$ is the set of all
choices of $M$ spins out of $N$.  We can then expand
\begin{align}
  X_{\m w} = \f1{M!} \sum_{\sigma\in\pi\p{M}} \sum_{k\in\C_N\p{M}}
  w_{\sigma\p{k}} X_{\sigma\p{k}}.
\end{align}
For simplicity, we will assume throughout this work that $\O$ is
homogeneous in the sense that $M$ is the same for all
$\p{X,\m w}\in\O$, with the understanding that a generalization to
different $M$ for each $\p{X,\m w}\in\O$ is straightforward.

@@

When the operator norm of $\hat\O$ is sufficiently small, namely
$\norm*{\hat\O}<\Delta_{\t{gap}}/2$, we can treat its effect on the PS
manifold $\M_0$ perturbatively.  The effective Hamiltonian
$H_{\t{eff}}$ induced by $\hat\O$ on the ground-state manifold $\M_0$
through second order in perturbation theory
is\cite{bravyi2011schrieffer, perlin2019effective}
\begin{align}
  H_{\t{eff}} = H_{\t{eff}}^{(1)} + H_{\t{eff}}^{(2)},
  &&
  H_{\t{eff}}^{(1)} = \P_0 \hat\O \P_0,
  &&
  H_{\t{eff}}^{(2)} = - \sum_{\Delta\ne0}
  \f1\Delta \P_0 \hat\O \P_\Delta \hat\O \P_0,
  \label{eq:H_eff}
\end{align}
where $\P_\Delta$ is a projector onto the eigenspace $\E_\Delta$ of
the interaction Hamiltonian $H_0$ with interaction energy $\Delta$
above that of PS manifold $\M_0$.  In particular, $\P_0$ is a
projector onto $\M_0$ itself, and $\E_0=\M_0$.  The first order
effective Hamiltonian $H_{\t{eff}}^{(1)}$ is straightforward to
simplify using the permutational symmetry of $\M_0$:
\begin{align}
  H_{\t{eff}}^{(1)} \EQPS \sum_{\p{X,\m w}\in\O} \bar w\,\col{X},
  \label{eq:H_eff_1}
\end{align}
where $\EQPS$ denotes equality under a restriction to $\M_0$; $\bar w$
is an average of $w_k$ over all $k\in\D_N\p{M}$; and $\col{X}$ is a
permutation-symmetrized collective version of $X$, i.e.
\begin{align}
  \bar w \equiv \f1{\abs{\D_N\p{M}}} \sum_{k\in\D_N\p{M}} w_k,
  &&
  \col{X} \equiv \f1{\abs{\pi\p{M}}} \sum_{\sigma\in\pi\p{M}}
  \sum_{k\in\C_N\p{M}} X_{\sigma\p{k}},
\end{align}
where, to be explicit,
\begin{align}
  \abs{\D_N\p{M}} = N!_M,
  &&
  \pi\p{M} = M!,
\end{align}
with $N!_M\equiv\prod_{j=0}^{M-1}\p{N-j}=M!\times{N\choose M}$
denoting a falling factorial.  An immediate consequence of
\eqref{eq:H_eff_1} is the fact that perturbing SU(2)-symmetric
interactions by two-body $s_\z\otimes s_\z$ interactions yields the
one-axis twisting (OAT) Hamiltonian\cite{kitagawa1993squeezed,
  ma2011quantum} $H_{\t{OAT}}=\chi S_\z^2$ at first order in
perturbation theory, with the OAT strength $\chi$ equal to the mean
coefficient of the two-body $s_\z\otimes s_\z$ terms.

@@

In this section, we simplify the products of the form
$H_0X_{\m w}\ket\psi$ to arrive at the expansion in
\eqref{eq:diagnosis}, which defines conditions under which a
multi-body perturbation $X_{\m w_\Delta}$ generates eigenstates of the
SU($n$)-symmetric interaction Hamiltonian $H_0$ in \eqref{eq:H_0} when
acting on a permutationally symmetric state $\ket\psi\in\M_0$.  Here
$X_{\m w}$ is an $M$-body perturbation defined by
\begin{align}
  X_{\m w} \equiv \f1{M!} \sum_{k\in\D_N\p{M}} w_k X_k,
  \label{eq:pert_copy}
\end{align}
where $X$ is an $M$-spin operator; $k\equiv\p{k_1,k_2,\cdots,k_M}$ is
a list of $M$ spins that $X_k$ acts on; $\D_N\p{M}$ is defined in
\eqref{eq:off_diags}; and $\m w$ is a dimension-$M$ (i.e.~$M$-index)
tensor with scalar entries $w_k$.

@@

Given a system of $N$ spins, we wish to project a product of
multi-body operators onto the permutationally symmetric (PS) manifold
$\M_0$.  For $p$ multi-body operators, such a product takes the form
\begin{align}
  \m X_{\m w}
  \equiv \prod_{\alpha\in\ZZ_p} \f1{M_\alpha!}
  \sum_{k_\alpha\in\D_N\p{M_\alpha}}
  w_\alpha\p{k_\alpha} X_\alpha\p{k_\alpha},
  \label{eq:sym_prod_start}
\end{align}
where $\m X\equiv\p{X_1,X_2,\cdots,X_p}$ is a vector of
$M_\alpha$-spin operators $X_\alpha$;
$\m w\equiv\p{w_1,w_2,\cdots,w_p}$ is a vector of dimension-$M_\alpha$
tensors $w_\alpha$;
$k_\alpha\equiv\p{k_{\alpha,1},k_{\alpha,2},\cdots,k_{\alpha,M_\alpha}}$
is a choice of $M_\alpha$ spins that $X_\alpha\p{k_\alpha}$ acts on;
$\D_N\p{M_\alpha}$ is the set of all ordered choices of $M_\alpha$
elements from a set of size $N$; and $w_\alpha\p{k_\alpha}$ is a
scalar entry of $w_\alpha$.  For convenience, we define
$w_\alpha\p{k_\alpha}$ and $X_\alpha\p{k_\alpha}$ for all
$k_\alpha\in\ZZ_N^{M_\alpha}$ by enforcing that
$w_\alpha\p{k_\alpha}=X_\alpha\p{k_\alpha}=0$ for all $k_\alpha$ with
repeated values.  These definitions allow us to replace sums over
$\D_N\p{M_\alpha}$ by sums over $\ZZ_N^{M_\alpha}$, thereby making the
sums over the individual indices in $k_\alpha$ independent.

To simplify the product in \eqref{eq:sym_prod_start}, we first replace
all $\D_N\p{M_\alpha}$ by $\ZZ_N^{M_\alpha}$ and rewrite the product
of sums as a sum of products:
\begin{align}
  \m X_{\m w}
  &= \f1{M_1!} \sum_{k_1\in\ZZ_N^{M_1}}
  \f1{M_2!} \sum_{k_2\in\ZZ_N^{M_1}} \cdots
  \f1{M_p!} \sum_{k_p\in\ZZ_N^{M_p}}
  \prod_{\alpha\in\ZZ_p} w_\alpha\p{k_\alpha} X_\alpha\p{k_\alpha} \\
  &= \f1{\m M!} \sum_{k\in\ZZ_N^{\m M}} \prod_{\alpha\in\ZZ_p}
  w_\alpha\p{k_\alpha} X_\alpha\p{k_\alpha},
  \label{eq:sum_of_products}
\end{align}
where $\m M\equiv\p{M_1,M_2,\cdots,M_p}$ is a vector of the dimensions
$M_\alpha$; $\m M!\equiv\prod_{M_\alpha\in\m M}M_\alpha!$ for
shorthand; and
$\ZZ_N^{\m M}\equiv\bigotimes_{\alpha\in\ZZ_p}\ZZ_N^{M_\alpha}$, such
that each $k\in\ZZ_N^{\m M}$ has the decomposition
$k=\p{k_1,k_2,\cdots,k_p}$ with $k_\alpha\in\ZZ_N^{M_\alpha}$.

We now classify terms in \eqref{eq:sum_of_products} by the numbers
$g_S$ of indices shared by all tensors $w_\alpha$ with
$\alpha\in S\subset\ZZ_p$.  For example, a term of the form
$w_1\p{a,b,c} w_2\p{b,d,e} w_3\p{b,c,d,e}$ with distinct indices
$a,b,c,d,e$ would have
\begin{align}
  g_{\set{1}} = 1,
  &&
  g_{\set{1,2,3}} = 1,
  &&
  g_{\set{1,3}} = 1,
  &&
  g_{\set{2,3}} = 2,
\end{align}
and $g_S=0$ for all other subsets $S\subset\ZZ_3$.  This index
assignment can be associated with the Venn diagram
\begin{align}
  \diagram{example_123},
  \label{eq:venn_diagram}
\end{align}
where $g_S$ is determined the number of dots at the intersection of
circles $\alpha\in S$.  We denote the set of all nonempty subsets of
$\ZZ_p$ by $\PP^*\p{\ZZ_p}$, and denote a choice of $g_S$ for all
$S\in\PP^*\p{\ZZ_p}$ by $g$.  With representations such as
\eqref{eq:venn_diagram} in mind, we will loosely refer to $g$ as
simply a ``diagram''.  We define
$\abs{g}\equiv\sum_{S\in\PP^*\p{\ZZ_p}}g_S$ to be the number of
indices considered in $g$ (or equivalently the number of dots in the
diagram $g$), and we denote the set of all diagrams for
$p=\dim\p{\m M}$ tensors with dimensions $\m M$ by $\G\p{\m M}$.  For
consistency, all diagrams $g\in\G\p{\m M}$ must associate exactly
$M_\alpha$ indices with the dimension-$M_\alpha$ tensor $w_\alpha$,
i.e.
\begin{align}
  \sum_{S\in\PP^*\p{\ZZ_p}\,:\,\alpha\in S} g_S
  = \sum_{R\in\PP^*\p{\ZZ_p\setminus\set{\alpha}}} g_{\set{\alpha}\cup R}
  = M_\alpha
\end{align}
for all $\alpha\in\ZZ_p$.

Schematically, a classification of the terms in
\eqref{eq:sum_of_products} by diagrams $g\in\G\p{\m M}$ will allow us
to expand
\begin{align}
  \m X_{\m w} \EQPS \sum_{g\in\G\p{\m M}} \m X_{\m w}\p{g},
  \label{eq:sym_prod_diagrams}
\end{align}
where $\EQPS$ denotes equality up to a restriction to the PS manifold
$\M_0$; and $\m X_{\m w}\p{g}$ is the sum of all terms in
\eqref{eq:sum_of_products} that are associated with the diagram $g$.
To collect these terms, we first consider all distinct ways to assign
$\abs{g}$ indices to specific tensor factors of $\m w$ and $\m X$ in a
manner consistent with $g$.  For example, if $\m X=\p{X_1,X_2}$ with
dimensions $\p{M_1,M_2}=\p{2,1}$ and $\abs{g}=2$ with
$g_{\set{1,2}}=g_{\set{2}}=1$, then $X_1\p{a,b} X_2\p{b}$ and
$X_1\p{b,a} X_2\p{b}$ are distinct assignments of $\abs{g}$ indices to
$\m X$ in a manner consistent with $g$, provided that $a\ne b$.  After
assigning $\abs{g}$ indices to $\m w$ and $\m X$ in a manner
consistent with $g$, we need to sum over all values that these indices
can take, which amounts to a sum over all
$k=\p{k_1,k_2,\cdots,k_{\abs{g}}}\in\ZZ_N^{\abs{g}}$ with the
restriction that $k_\mu\ne k_\nu$ for all $\mu\ne\nu$ to enforce that
all indices are distinct, or equivalently a sum over all
$k\in\D_N\p{\abs{g}}$.

>>>>

\newcommand{\A}{\mathcal{A}}
\newcommand{\B}{\mathcal{B}}
\newcommand{\C}{\mathcal{C}}
\newcommand{\D}{\mathcal{D}}
\newcommand{\E}{\mathcal{E}}
\newcommand{\G}{\mathcal{G}}
\newcommand{\I}{\mathcal{I}}
\renewcommand{\L}{\mathcal{L}}
\newcommand{\M}{\mathcal{M}}
\newcommand{\N}{\mathcal{N}}
\renewcommand{\O}{\mathcal{O}}
\renewcommand{\P}{\mathcal{P}}
\renewcommand{\S}{\mathcal{S}}
\newcommand{\T}{\mathcal{T}}

@@

We consider an array of $N$ multilevel spins with SU($n$)-symmetric
interactions that can be written in the form
\begin{align}
  H_0 = \sum_{\substack{p,q\in\ZZ_N\\p<q}} h_{pq} \Pi_{pq},
  &&
  \Pi_{pq} \equiv \sum_{\mu,\nu\in\ZZ_n} s_{\mu\nu}^{(p)} s_{\nu\mu}^{(q)},
  \label{eq:H_0}
\end{align}
where $p,q$ index individual spins; $\ZZ_k$ is the set of integers
modulo $k$; $h_{pq}$ are scalar coupling constants; $\mu,\nu$ index
states in an orthonormal basis for the $n$-dimensional Hilbert space
of a single spin; the operator $s_{\mu\nu}^{(p)}\equiv\op{\mu}{\nu}_p$
flips the state of spin $p$ to $\ket\nu$ from $\ket\mu$; and the
operator $\Pi_{pq}$ permutes the states of spins $p$ and $q$.  The
permutation operator $\Pi_{pq}$ is essentially a multilevel
generalization of the SU(2)-symmetric spin interaction
$\v s_p\c\v s_q$, with $\v s\equiv\p{\v s_\x,s_\y,s_\z}$,
$s_\alpha\equiv\sigma_\alpha/2$, and $\sigma_\alpha$ a single-spin
Pauli operator.

If the coefficients $h_{pq}$ of the interaction Hamiltonian $H_0$ are
all negative, then the ground-state manifold $\M_0$ of the interaction
Hamiltonian $H_0$ consists of permutationally symmetric (PS) states
that are simultaneous $+1$ eigenstates of all permutation operators
$\Pi_{pq}$.  We will assume that all $h_{pq}<0$ throughout these
notes, but comment that this restriction can be relaxed to the
assumption that the initial state of all spins is permutationally
symmetric (e.g.~a spin-polarized state).  We will also assume that the
PS manifold $\M_0$ is gapped away from all orthogonal states by an
interaction energy $\Delta_{\t{gap}}>0$.  Power-law couplings of the
form $h_{pq}\sim-1/\abs{r_p-r_q}^\alpha$ with positions $r_p,r_q$ on a
$D$-dimensional lattice, for example, always yield a non-vanishing
spectral gap $\Delta_{\t{gap}}$ for long-range interactions with
$\alpha\le D$ (see Appendix \ref{sec:spectral_gap}).

The PS manifold $\M_0$ is spanned by a basis of states
$\ket{m}=\ket{m_1,m_2,\cdots,m_n}$ with a definite occupation number
$m_\mu$ of the single-spin state $\mu$, and $\sum_\mu m_\mu=N$.  In
the case of $n=2$, for example, the PS manifold $\M_0$ is precisely
the Dicke manifold\cite{dicke1954coherence} of $N$ qubits, spanned by
Dicke states $\set{\ket{m_0,m_1}}$ for which $m_0+m_1=N$ and the
integer $m_0$ ($m_1$) indicates the total occupation number of qubit
state $\ket{0}$ ($\ket{1}$).  The dimension of the PS manifold $\M_0$
is determined by the number of ways to assign $N$ spins to $n$ states,
which is ${N+n-1 \choose n-1}\sim N^{n-1}$.  The energy of any PS
state $\ket\psi\in\M_0$ with respect to the interaction Hamiltonian
$H_0$ in \eqref{eq:H_0} is
\begin{align}
  E_0 = \sum_{\substack{p,q\in\ZZ_N\\p<q}} h_{pq}.
\end{align}

@@

In order to induce non-trivial dynamics for initially PS states
$\ket\psi\in\M_0$, which are otherwise stationary eigenstates of the
interaction Hamiltonian $H_0$ in \eqref{eq:H_0}, we will add to the
Hamiltonian an operator $\hat\O$ of the general form
\begin{align}
  \hat\O \equiv \sum_{\p{X,\m w}\in\O} X_{\m w},
  &&
  X_{\m w} \f1M! \equiv \sum_{k\in\D_N\p{M}} w_k X_k,
  \label{eq:pert}
\end{align}
where $\hat\O$ is constructed from the set $\O\equiv\set{\p{X,\m w}}$
of $M$-spin operators $X$ together with dimension-$M$ (i.e.~$M$-index)
tensors $\m w$; $k=\p{k_1,k_2,\cdots,k_M}$ is a list of of $M$ spins
that $X_k$ acts on; $w_k$ is a scalar entry of $\m w$; and
\begin{align}
  \D_N\p{M} \equiv
  \set{ k \in \ZZ_N^M: \t{all entries of}~k~\t{are distinct} },
  \label{eq:off_diags}
\end{align}
is the strictly off-diagonal part of $\ZZ_N^M$.  The factor of $1/M!$
is included in \eqref{eq:pert} by convention to average over the $M!$
terms addressing any given set of spins.  Due to the permutational
symmetry enforced by the interaction Hamiltonian $H_0$, we will find
it convenient to decompose $\D_N\p{M}=\pi\p{M}\times\C_N\p{M}$, where
$\pi\p{M}$ is the permutation group of $\ZZ_M$, and
\begin{align}
  \C_N\p{M} \equiv \set{ k \subset \ZZ_N: \abs{k} = M },
  \label{eq:choices}
\end{align}
is the set of subsets (``choices'') of $M$ elements from $\ZZ_N$.  We
can then expand
\begin{align}
  X_{\m w} = \f1{M!} \sum_{\sigma\in\pi\p{M}} \sum_{k\in\C_N\p{M}}
  w_{\sigma\p{k}} X_{\sigma\p{k}}.
\end{align}
For simplicity, we will assume throughout this work that $\hat\O$ is
homogeneous in the sense that $M$ is the same for all
$\p{X,\m w}\in\O$, with the understanding that a generalization to
inhomogeneous operators $\hat\O$ is straightforward.

@@

When the operator norm of $\hat\O$ is sufficiently small, namely
$\norm*{\hat\O}<\Delta_{\t{gap}}/2$, we can treat its effect on the PS
manifold $\M_0$ perturbatively.  The effective Hamiltonian
$H_{\t{eff}}$ induced by $\hat\O$ on the ground-state manifold $\M_0$
through second order in perturbation theory
is\cite{bravyi2011schrieffer, perlin2019effective}
\begin{align}
  H_{\t{eff}} = H_{\t{eff}}^{(1)} + H_{\t{eff}}^{(2)},
  &&
  H_{\t{eff}}^{(1)} = \P_0 \hat\O \P_0,
  &&
  H_{\t{eff}}^{(2)} = - \sum_{\Delta\ne0}
  \f1\Delta \P_0 \hat\O \P_\Delta \hat\O \P_0,
  \label{eq:H_eff}
\end{align}
where $\P_\Delta$ is a projector onto the eigenspace $\E_\Delta$ of
the interaction Hamiltonian $H_0$ with interaction energy $\Delta$
above that of PS manifold $\M_0$.  In particular, $\P_0$ is a
projector onto $\M_0$ itself, and $\E_0=\M_0$.  The first order
effective Hamiltonian $H_{\t{eff}}^{(1)}$ is straightforward to
simplify using the permutational symmetry of $\M_0$:
\begin{align}
  H_{\t{eff}}^{(1)} \EQPS \sum_{\p{X,\m w}\in\O} \bar w\,\col{X},
  \label{eq:H_eff_1}
\end{align}
where $\EQPS$ denotes equality under a restriction to $\M_0$; the
coefficient $\bar w$ is the average of all $w_k$; and the collective
operator $\col{X}$ is a permutation-averaged sum of all $X_k$, i.e.
\begin{align}
  \bar w \equiv \f1{\abs{\D_N\p{M}}}
  \sum_{k\in\D_N\p{M}} w_{\sigma\p{k}},
  &&
  \col{X} \equiv \f1{\abs{\pi\p{M}}} \sum_{\sigma\in\pi\p{M}}
  \sum_{k\in\C_N\p{M}} X_{\sigma\p{k}},
\end{align}
with $\abs{\D_N\p{M}}=M!\times{N\choose M}=\prod_{j=0}^{M-1}\p{N-j}$
and $\abs{\pi\p{M}}=M!$.  An immediate consequence of
\eqref{eq:H_eff_1} is the fact that perturbing SU(2)-symmetric
interactions by two-body $s_\z\otimes s_\z$ interactions yields the
one-axis twisting (OAT) Hamiltonian\cite{kitagawa1993squeezed,
  ma2011quantum} $H_{\t{OAT}}=\chi S_\z^2$ at first order in
perturbation theory, with the OAT strength $\chi$ equal to the mean
coefficient of the two-body $s_\z\otimes s_\z$ terms.

@@

In this section, we simplify the products of the form
$H_0X_{\m w}\ket\psi$ to arrive at the expansion in
\eqref{eq:diagnosis}, which defines conditions under which a
multi-body perturbation $X_{\m w_\Delta}$ generates eigenstates of the
SU($n$)-symmetric interaction Hamiltonian $H_0$ in \eqref{eq:H_0} when
acting on a permutationally symmetric state $\ket\psi\in\M_0$.  Here
$X_{\m w}$ is an $M$-body perturbation defined by
\begin{align}
  X_{\m w}
  \equiv \sum_{k\in\D_N\p{M}} w_k X_k
  = \sum_{k\in\C_N\p{M}} \sum_{\sigma\in\pi\p{M}}
  w_{\sigma\p{k}} X_{\sigma\p{k}},
\end{align}
where $X$ is an $M$-spin operator; $k\equiv\p{k_1,k_2,\cdots,k_M}$ is
a list of $M$ spins that $X_k$ acts on; $\m w$ is a dimension-$M$
(i.e.~$M$-index) tensor with scalar entries $w_k$; $\D_N\p{M}$ is the
strictly off-diagonal part of $\ZZ_N^M$, defined in
\eqref{eq:off_diags}; $\C_N\p{M}$ is the set of all choices of $M$
elements from $\ZZ_N$, defined in \eqref{eq:choices}; and $\pi\p{M}$
is the permutation group of $\ZZ_M$.

@@

Given a system of $N$ spins, we wish to project a product of
multi-body operators onto the permutationally symmetric (PS) manifold
$\M_0$.  For $p$ multi-body operators, such a product takes the form
\begin{align}
  \m X_{\m w}
  \equiv \prod_{\alpha\in\ZZ_p} \f1{M_\alpha!}
  \sum_{k_\alpha\in\D_N\p{M_\alpha}}
  w_\alpha\p{k_\alpha} X_\alpha\p{k_\alpha}
  = \f1{\m M!} \sum_{k\in\D_N\p{\m M}} \prod_{\alpha\in\ZZ_p}
  w_\alpha\p{k_\alpha} X_\alpha\p{k_\alpha},
  \label{eq:sym_prod_start}
\end{align}
where $\m X\equiv\p{X_1,X_2,\cdots,X_p}$ is a vector of
$M_\alpha$-spin operators $X_\alpha$;
$\m w\equiv\p{w_1,w_2,\cdots,w_p}$ is a vector of dimension-$M_\alpha$
tensors $w_\alpha$; $\m M\equiv\p{M_1,M_2,\cdots,M_p}$ is a vector of
the dimensions $M_\alpha$;
$k_\alpha=\p{k_{\alpha,1},k_{\alpha,2},\cdots,k_{\alpha,M_\alpha}}$ is
a choice of $M_\alpha$ spins that $X_\alpha\p{k_\alpha}$ acts on;
$w_\alpha\p{k_\alpha}$ is a scalar entry of $w_\alpha$;
$\D_N\p{M_\alpha}$ is the strictly off-diagonal part of $\ZZ_N^M$,
defined in \eqref{eq:off_diags};
$\D_N\p{\m M}\equiv\bigotimes_{M_\alpha\in\m M}\D_N\p{M_\alpha}$, such
that each $k\in\D_N\p{\m M}$ has the decomposition
$k=\p{k_1,k_2,\cdots,k_p}$ with $k_\alpha\in\D_N\p{M_\alpha}$; and
$\m M!\equiv\prod_{M_\alpha\in\m M}M_\alpha!$ for shorthand.  For
later convenience, we define
$w_\alpha\p{k_\alpha}=X_\alpha\p{k_\alpha}=0$ for any $k_\alpha$ with
repeated entries.

We now classify terms in \eqref{eq:sym_prod_start} by the numbers
$g_S$ of indices shared by all tensors $w_\alpha$ with
$\alpha\in S\subset\ZZ_p$.  For example, a term of the form
$w_1\p{a,b,c} w_2\p{b,d,e} w_3\p{b,c,d,e}$ with distinct indices
$a,b,c,d,e$ would have
\begin{align}
  g_{\set{1}} = 1,
  &&
  g_{\set{1,2,3}} = 1,
  &&
  g_{\set{1,3}} = 1,
  &&
  g_{\set{2,3}} = 2,
\end{align}
and $g_S=0$ for all other subsets $S\subset\ZZ_3$.  This index
assignment can be associated with the Venn diagram
\begin{align}
  \diagram{example_123},
  \label{eq:venn_diagram}
\end{align}
where $g_S$ is determined the number of dots at the intersection of
circles $\alpha\in S$.  We denote the set of all nonempty subsets of
$\ZZ_p$ by $\PP^*\p{\ZZ_p}$, and denote a choice of $g_S$ for all
$S\in\PP^*\p{\ZZ_p}$ by $g$.  With representations such as
\eqref{eq:venn_diagram} in mind, we will loosely refer to $g$ as
simply a ``diagram''.  We define
$\abs{g}\equiv\sum_{S\in\PP^*\p{\ZZ_p}}g_S$ to be the number of
indices considered in $g$ (or equivalently the number of dots in the
diagram $g$), and we denote the set of all diagrams for
$p=\dim\p{\m M}$ tensors with dimensions $\m M$ by $\G\p{\m M}$.  For
consistency, all diagrams $g\in\G\p{\m M}$ must associate exactly
$M_\alpha$ indices with the dimension-$M_\alpha$ tensor $w_\alpha$,
i.e.
\begin{align}
  \sum_{S\in\PP^*\p{\ZZ_p}\,:\,\alpha\in S} g_S
  = \sum_{R\in\PP^*\p{\ZZ_p\setminus\set{\alpha}}} g_{\set{\alpha}\cup R}
  = M_\alpha
\end{align}
for all $\alpha\in\ZZ_p$.  A classification of the terms in
\eqref{eq:sym_prod_start} by diagrams $g\in\G\p{\m M}$ allows us to
expand
\begin{align}
  \m X_{\m w} \EQPS \sum_{g\in\G\p{\m M}} \m X_{\m w}\p{g},
  \label{eq:sym_prod_diagrams}
\end{align}
where $\EQPS$ denotes equality under a restriction to $\M_0$, and
$\m X_{\m w}\p{g}$ is the sum of all terms in
\eqref{eq:sym_prod_start} that are associated with the diagram $g$.

To find an explicit form of $\m X_{\m w}\p{g}$, we first consider all
distinct ways to assign $\abs{g}$ indices to specific tensor factors
of $\m w$ and $\m X$ in a manner consistent with $g$.  For example, if
$\m X=\p{X_1,X_2}$ with dimensions $\p{M_1,M_2}=\p{2,1}$ and
$\abs{g}=2$ with $g_{\set{1,2}}=g_{\set{2}}=1$, then
$X_1\p{a,b} X_2\p{b}$ and $X_1\p{b,a} X_2\p{b}$ are distinct
assignments of $\abs{g}$ indices to $\m X$ in a manner consistent with
$g$, provided that $a\ne b$.  After assigning $\abs{g}$ indices to
$\m w$ and $\m X$ in a manner consistent with $g$, we need to sum over
all values that these indices can take, which amounts to a sum over
all $k\in\D_N\p{\abs{g}}$.
